Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-10-21 Cost_manuscript.Rmd
gathering writing samples for the intro

  Effort is costly, but sometimes it can add value (For the talk, use the T example with Angelina, but think about how well it fits in the "prefer effort"). In thinking about this, studies have focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. (other frame: What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands). Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

@@

In this talk, I'll show you how we'veused this framework to understand the inherent (elaborate on inherent above) and relative costly nature of effort and delay. I'll focus on behavioral findings, and I'll give a nudge to modeling. Then I'll discuss how we plan to understand the brain networks that characterize this behavior, with a focus on individual differences.

What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands.

- I guess you can mention that the wait group acts as a baseline of foraging behavior.
- Re-read Mikey's effort value paper to see if there is anything there that can allow us to transition into foraging. Like, are there elements that make cog effort valuable that are well encapsuled by foraging paradigms?
- Pending questions: explain logistic coefficients in exponentiated form? (i.e. X% more/less likely than...)

@@

- Make sure to note that we controlled for time as a confound for effort, as more effortful tasks normally take longer
- What we think: subjective timing is shortened by the way we implemented cognitive effort. This can be recalibrated if many task types are experienced. Maybe the implementation allowed for counting. Alternatively, working memory engagement could produce timing issues (Buhi n Gershman). Also look at what Andrew Westbrook shared. Add this to a potential fatigue, although that would also mean that waiting would be accepted less over time in exp 1.
- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people)
- Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effor aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort
- It's true that we didnt calibrate cog effort per subject. But instead we brought all subjects to the same level of performance, and we saw that as time progressed they became less likely to accept longer periods of effort.

>>>>

  Prospectus: Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. *Effort examples here?* In neuroeconomics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions. For example, we might choose to take the effortful route home that we normally avoid rather than waiting for a delayed train. Understanding the sporadic attractiveness of effort exemplified above has been elusive, and presents an important research avenue (Inzlicht, Shenhav, & Olivola, 2017). 
  Measuring the cost of effort and delay is often done by offering higher payouts for an increment of the respective demand, but flexible preferences for effort might stem from the availability of other types of demands in the same choice environment. (Kool & Botvinick, 2018). Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).
  In this study, we will test the hypothesis that preferences will vary depending on whether individuals face a single or multiple types of behavioral demands, and identify what factors drive these choices. Based on two foraging experiments and computational modeling, we will show that 1) preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.
  In this study, we will behaviorally and computationally test the hypothesis that the attractiveness of a reward that requires effort or delay depends on whether rewards associated with alternative demands are available in the same context. Using a foraging paradigm will enable us to 1) assess whether people integrate the statistics of the environment to generate strategies that maximize their earnings; 2) compare how the presence of effort or delay shifts such strategies, as reflected by acceptance rates; 3) examine if acceptance rates are further altered by foraging for costs in addition to rewards; and 4) test a combination of foraging (Constantino & Daw, 2015; Stephens & Krebs, 2019; Fawcett, McNamara, & Houston, 2012) and reinforcement learning models (Sutton & Barto, 2018) to examine whether preferences for demands reflect a subjective reduction of cost or boost of value, and determine any factors besides contextual manipulations that participants consider in order to adjust their choices.

  Grant: Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. Demands for both time and effort have been identified as costs that detract from the net subjective value of rewards. Decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [1–9]. Excessive aversion to delay and effort can manifest as impulsivity and apathy, respectively [1,10–12], fluctuations in which are hallmark symptoms of various mental illnesses and substance use disorders [13]. However, we do not yet fully understand 1) how the brain transforms these two parallel forms of discounting into seemingly opposite behaviors; 2) whether the same mechanism applies to different types of effort (e.g. physical versus cognitive); or 3) the ways in which these distinct forms of cost interact in individuals. In this project, I plan to combine individualized brain mapping with computational modeling to examine how an individual’s functional brain network organization relates to the subjective costs of physical effort, cognitive effort, and pure delay.
  Behavioral distinctions between effort and delay costs: The process of engaging with a costly reward can be divided into three stages: evaluation (when costs and benefits of a prospect are integrated), pursuit (i.e. waiting or exerting effort), and receipt of the reward. An ever-increasing literature in psychology, neuroscience, and economics has focused on the first stage to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated demand [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. Investigations of the pursuit and receipt stages have been comparatively limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [5,11,15,16,19]. However, discrepancies between expected and experienced value are well known [20], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards [3,21,22].
  The evidence reviewed above has two main limitations. First, studies have not focused on how effort and delay might interact in the common case of choosing among demands. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. 
  It is unknown how effort demands discount value in isolation versus when alternative demands are available. The friction that we perceive from these demands could depend on what we could be performing instead [61].

IDEAS:

   Studies have often focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. (other frame: What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands). Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

@@

  Modeling: mention that the rich history of foraging modeling can help us pinpoint a number of things, like the putative cost or attractiveness of a demand, whether preferences emerge from focal considerations of the value of a trial or a global estimate of what alternative way could be used to obtain rewards. Foraging and its modeling can provide a common framework to examine differences, rather than relying on matching disc functions.

  Clinical: Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- I guess you can mention that the wait group acts as a baseline of foraging behavior.
- Re-read Mikey's effort value paper to see if there is anything there that can allow us to transition into foraging. Like, are there elements that make cog effort valuable that are well encapsuled by foraging paradigms?
- Pending questions: explain logistic coefficients in exponentiated form? (i.e. X% more/less likely than...)

@@

- Make sure to note that we controlled for time as a confound for effort, as more effortful tasks normally take longer
- What we think: subjective timing is shortened by the way we implemented cognitive effort. This can be recalibrated if many task types are experienced. Maybe the implementation allowed for counting. Alternatively, working memory engagement could produce timing issues (Buhi n Gershman). Also look at what Andrew Westbrook shared. Add this to a potential fatigue, although that would also mean that waiting would be accepted less over time in exp 1.
- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people)
- Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort
- It's true that we didnt calibrate cog effort per subject. But instead we brought all subjects to the same level of performance, and we saw that as time progressed they became less likely to accept longer periods of effort.
